Mental Health Plays Significant Role in Physical Wellness From 1972 to 2022 US General Social Surveys

### 2024-04-23 paper/paper.qmd
fix result section

To speculate possible demographic factors influencing days of feeling mentally and physically unwell in a month, gender and sex are used. In general, females from 20 to 39 years old are more likely to report a higher number of days feeling mentally unwell (@fig-demo-mental). In terms of physical health, females from 40 to 59 years old report a higher number of days of feeling unwell (@fig-demo-physical).

As we can see, females in general report a higher number of counted days of feeling both physically and mentally unwellness in a month. The difference could be discrimination between gender groups, such as females usually being less represented in medical research. On the other hand, females are reported to be more sensitive to their mental health, which could be the reason why they tend to have a higher number of reports of being mentally unwell.

@@

@@

In addition to model and demographic analyses, a variable indicating respondents' general health was as well included for supplementary support of the paper. The general health variable is categorical, consisting of four responses: "1" for Excellent, "2" for Good, "3" for Fair, and "4" for Poor. As shown in @fig-mental-vs-general, we can see that the number of days feeling mentally unwell has a positive correlation with the general health status, meaning that if an individual has more days feeling mentally unwell, they are more likely to have poor general health as well. This result connects back to our estimand, supporting mental health's influence on one's physical health status.

@@

Based on @tbl-modelresults, however, although having influences on linear equation in @fig-mental-vs-physical, depression diagnosis's influences on counted days of feeling physically unwell is not significant ($\beta_2$ = −1.43, p = .22). Counted days of feeling mentally unwell, on the other hand, do have a significant positive correlation with counted days of feeling physically unwell ($\beta_1$ = 0.27, p = .01). Even though one of the correlation is not significant, it is still important for us to consider the variable for possible future research.

@@

## Weaknesses and Next Steps

@@

@@

>>>>

To speculate possible demographic factors influencing days of feeling mentally and physically unwell in a month, gender and sex are used. In general, most of the participants reported having less than 10 days of feeling unwell both mentally and physically. Throughout the responses except for 0 days, the occurrence of responses from female participants is higher than from male participants; this means that a larger percentage of female participants have experienced both mentally and physically unwell in the past 30 days. Age cohorts from 20 to 59 contribute to the majority of the occurrences for each response, which could be a result of the fact that the age cohort takes up to 87% of the dataset. In conclusion, females from 20 to 39 years old are more likely to report a higher number of days feeling mentally unwell (@fig-demo-mental). In terms of physical health, females from 40 to 59 years old report a higher number of days of feeling unwell (@fig-demo-physical).

As we can see, females in general report a higher number of counted days of feeling both physically and mentally unwellness in a month. The difference could be disparity occurring between gender groups, such as females usually being less represented in medical research. On the other hand, females are reported to be more sensitive to their mental health, which could be the reason why they tend to have a higher number of reports of being mentally unwell.

@@

## Nonresponse Rate

As we illustrated in @fig-nonresponse-rate, non-response rates may also help indicate significant differences in participants' attitudes toward each question included in the survey. Physical health-related questions have the highest non-response rate, indicating that most of the participants lack the motivation to time on speculate their physical health. On the other hand, the non-response rate for mental health and depression diagnosis is lower than for physical health questions. This difference between non-response rates could show that participants are more likely to pay attention when they are feeling mentally unwell and are more willing to answer questions corresponding to mental health.

@@

In addition to model and demographic analyses, a variable indicating respondents' general health was also included for supplementary support of the paper. The general health variable is categorical, consisting of four responses: "1" for Excellent, "2" for Good, "3" for Fair, and "4" for Poor. As shown in @fig-mental-vs-general, we can see that the number of days feeling mentally unwell has a positive correlation with the general health status, meaning that if an individual has more days feeling mentally unwell, they are more likely to have poor general health as well. This result connects back to our estimand, supporting mental health's influence on one's physical health status. The relationship, however, is not strong ($\beta_1 =0.03$, $R^2 = 0.07$); this could be due to the fact that the health variable was not framed to be correlated to mental health. The variable may fail to associate with other health variables.

@@

Based on @tbl-modelresults, however, although having influences on linear equation in @fig-mental-vs-physical, depression diagnosis's influences on counted days of feeling physically unwell is not significant ($\beta_2$ = −1.43, p = .22). Counted days of feeling mentally unwell, on the other hand, do have a significant positive correlation with counted days of feeling physically unwell ($\beta_1$ = 0.27, p = .01). From @fig-mental-vs-physical, it is evident to state that depression diagnosis may be a moderator to strengthen the relationship between counted ays of mental and physical unwellness in the past 30 days. The lack of significance could be due to possible errors, validity of survey questions, biases, or other possible factors. Hence, even though one of the correlations is not significant, it is still important for us to consider the variable for possible future research.

@@

## Weaknesses

@@

### Model Limitations

@@

## Future Directions
